---
title: "The Dublin mess"
author: "Duc-Quang Nguyen"
date: "25 Oct 2016"
output: 
  html_document:
    toc: true
    toc_depth: 3
    theme: simplex
---

```{r, message = F, include = F, warning = F}
library(eurostat)
library(dplyr)
library(tidyr)
library(magrittr)
library(countrycode)
library(ggplot2)
library(scales)
library(swiTheme)
```

## Get Data
### Eurostat data via API
```{r getData, include = F}

downloadData <- F
saveMergedData <- "input/"


tables <- c('migr_asyappctza', 'migr_dubro', 'migr_dubri', 'migr_dubdi', 'migr_dubto', "migr_dubredaco", 'migr_dubdo', 'migr_dubdedaco')
names(tables) <- c(
  "data/asylumYearly.csv", "data/dublinOutR.csv",  "data/dublinInR.csv", "data/decisionDublinIn.csv", "data/dublinOutT.csv", 
  "data/dublinOutREurodac", "data/decisionDublinOut.csv", "data/dublinOutEurodacDecision")

if(downloadData) {
  # warning the data to download is very large over 1 Gb !
  getData <- function(dataID) {
  cat("\n", dataID)
   dat <- get_eurostat(dataID, time_format = "date", cache = F)
  }
  data.dl <- lapply(tables, getData)
  
  # get yearly asylum applications
  asyl <- data.dl[[1]] %>% filter(sex == "T", age == "TOTAL", asyl_app == "ASY_APP") %>% 
    select(one_of(c("citizen", "geo", "time", "values")))
  write.csv(asyl, file = names(tables)[1], row.names = F)
  
  # get yearly Dublin outgoing requests
  outr <- data.dl[[2]] %>% filter(request == "REQ", leg_prov == "TOTAL") %>%
    select(one_of(c("partner", "geo", "time", "values")))
  write.csv(outr, file = names(tables)[2], row.names = F)
 
  # get yearly Dublin incoming requests
  inr <- data.dl[[3]] %>% filter(request == "REQ", leg_prov == "TOTAL") %>%
    select(one_of(c("partner", "geo", "time", "values")))
  write.csv(inr, file = names(tables)[3], row.names = F)
  
  # get yearly Dublin incoming requests' decisions
  inrd <- data.dl[[4]] %>%  filter(request == "REQ", leg_prov == "TOTAL") %>%
    select(one_of(c("partner", "geo", "time", "values", "decision")))
  write.csv(inrd, file = names(tables)[4], row.names = F)
  
  # get yearly Dublin outgoing transfers
  outt <- data.dl[[5]] %>% filter( leg_prov == "TOTAL", duration == "TOTAL") %>%
    select(one_of(c("partner", "geo", "time", "values")))
  write.csv(outt, file = names(tables)[5], row.names = F) 
  
  # get yearly Dublin outgoing EURODAC requests 
  outreu <- data.dl[[6]] %>% filter(request == "REQ", leg_prov == "EDAC") %>%
    select(one_of(c("partner", "geo", "time", "values")))
  write.csv(outreu, file = names(tables)[6], row.names = F)
  
    # get yearlyDublin outgoing requests' decisions
  outrd <- data.dl[[7]] %>% filter(request == "REQ", leg_prov == "TOTAL") %>%
    select(one_of(c("partner", "geo", "time", "values", "decision")))
  write.csv(outrd, file = names(tables)[7], row.names = F)
  
  # get decisions on outgoing Dublin based on EURODAC
  outreud <- data.dl[[8]] %>% filter(request == "REQ", leg_prov == "EDAC") %>%
    select(one_of(c("partner", "geo", "time", "values", "decision")))
  write.csv(outreud, file = names(tables)[8], row.names = F)
  
} else {
  data.read <- sapply(1:length(tables), function(i)
    read.csv(names(tables)[i], stringsAsFactors = FALSE, check.names = F))
  names(data.read) <-  gsub("\\.csv", "", gsub("data\\/", "", names(tables)))
}

```

### Shape the data
```{r shapeData, include=T}

iso2.subset <- c("DE", "CH", "FR", "IT", "AT", "SE", "UK", "ES", "PT", "EL", "HU")
names(iso2.subset) <- countrycode(iso2.subset, "iso2c", "country.name")
names(iso2.subset)[which(iso2.subset == "UK")] <- "United Kingdom"
names(iso2.subset)[which(iso2.subset == "EL")] <- "Greece"
stopifnot(!any(is.na(names(iso2.subset))))

# % missing data
missingData <- function(df) {
  round(tapply(df$values, df$time, function(v) (sum(is.na(v))/length(v)) ) * 100)
}
lapply(data.read, missingData)

# Define time limits 
time.limit <- range(data.read[[1]]$time) #as.Date(c("2009-01-01", "2014-01-01"))

filterByIso2 <- function(
  dd, citizen.filter = "TOTAL", geo.subset = iso2.subset, tStart = time.limit[1], tEnd = time.limit[2]
) {
 if("partner" %in% colnames(dd)) {
   dd %>% filter(partner == citizen.filter, geo %in% geo.subset, time >= tStart, time <= tEnd)     
 } else {
   dd %>% filter(citizen == citizen.filter, geo %in% geo.subset, time >= tStart, time <= tEnd)     
 }
}

data.sub <- sapply(data.read, filterByIso2)
data.all <- sapply(data.read, function(dd) filterByIso2(dd, geo.subset = unique(data.read[[2]]$geo)))

```

## Combine datasets

```{r comine datasets, echo = F, message = F}
# massive check!
stopifnot(nrow(data.sub[[1]]) == nrow(data.sub[[2]]))
stopifnot(nrow(data.sub[[1]]) == nrow(data.sub[[3]]))
stopifnot(nrow(data.sub[[1]]) == nrow(data.sub[['decisionDublinIn']]) / 3)
stopifnot(nrow(data.sub[[1]]) == nrow(data.sub[['decisionDublinOut']]) / 3)
stopifnot(nrow(data.sub[[1]]) == nrow(data.sub[['dublinOutEurodacDecision']]) / 3)
stopifnot(nrow(data.sub[[1]]) == nrow(data.sub[['dublinOutT']]))
stopifnot(nrow(data.sub[[1]]) == nrow(data.sub[['dublinOutREurodac']]))

#compute % positive decision rate
positiveDecision <- function(dd, name = "" ) {
  result <- dd %>% group_by(geo, time) %>%
    summarise(rate = values[which(decision == 'ACC')] / values[which(decision == 'TOTAL')]) %>% 
    ungroup()
  colnames(result)[which(colnames(result) == "rate")] <- name
  result
}

data.sub[['decisionDublinIn']] <- positiveDecision(data.sub[['decisionDublinIn']], 'incomingAccRequestRate')
data.sub[['decisionDublinOut']] <- positiveDecision(data.sub[['decisionDublinOut']], 'outgoingAccRequestRate')
data.sub[['dublinOutEurodacDecision']] <- positiveDecision(data.sub[['dublinOutEurodacDecision']], 'outgoingEuroAccRequestRate')

# join tables
data <- select(data.sub[['asylumYearly']], -citizen) %>% 
  left_join(data.sub[['dublinOutR']] %>% select(geo, values, time) %>% rename(outgoingRequest = values)) %>%
  left_join(data.sub[['dublinInR']] %>% select(geo, values, time) %>% rename(incomingRequest = values)) %>%
  left_join(data.sub[['dublinOutT']] %>% select(geo, values, time) %>% rename(outgoingTransfer = values)) %>%
  left_join(data.sub[['dublinOutREurodac']] %>% select(geo, values, time) %>% rename(outgoingEurodacRequest = values)) %>%
  left_join(data.sub[['decisionDublinIn']]) %>%
  left_join(data.sub[['decisionDublinOut']]) %>%
  left_join(data.sub[['dublinOutEurodacDecision']])

data <- data %>% 
  group_by(time, geo) %>% mutate(percOutgoingR = outgoingRequest / values) %>%
  group_by(time, geo) %>% mutate(percOutgoingT = outgoingTransfer / values) %>% 
  group_by(time, geo) %>% mutate(percEurodacOutgoing = outgoingEurodacRequest / outgoingRequest) %>%  
  ungroup()

data$time <- as.Date(data$time)
data$country <- names(iso2.subset)[match(data$geo, iso2.subset)]

vars <- colnames(data)[!colnames(data) %in% c('geo', 'time', 'country')]
names(vars) <- c(
  "Total number of asylum requests",
  "Total number of outgoing Dublin requests",
  "Total number of incoming Dublin requests",  
  "Total number of outgoing Dublin transfers",
  "Total number of outgoing Eurodac requests",
  "Acceptance rate of incoming Dublin requests",
  "Acceptance rate of outgoing Dublin requests",  
  "Acceptance rate of outgoing Eurodac requests",  
  "% of outgoing Dublin requests out of all asylum requests",
  "% of outgoing Dublin transfers out of all asylum requests",  
  "% of Eurodac requests out of all outgoing requests")

```

```{r plot total over all countries, message = F, warning = F, echo = F}

positiveDecision2 <- function(dd, name = "" ) {
  result <- dd %>% group_by(time) %>%
    summarise(rate = sum(values[which(decision == 'ACC')], na.rm = T) / sum(values[which(decision == 'TOTAL')], na.rm = T)) %>% 
    ungroup()
  colnames(result)[which(colnames(result) == "rate")] <- name
  result
}

data.all[['decisionDublinIn']] <- positiveDecision2(data.all[['decisionDublinIn']], 'incomingAccRequestRate')
data.all[['decisionDublinOut']] <- positiveDecision2(data.all[['decisionDublinOut']], 'outgoingAccRequestRate')
data.all[['dublinOutEurodacDecision']] <- positiveDecision2(data.all[['dublinOutEurodacDecision']], 'outgoingEuroAccRequestRate')

data.all <- data.all[[1]] %>% group_by(time) %>% summarise(values = sum (values, na.rm = T))  %>% 
  left_join(data.all[['dublinOutR']] %>% group_by(time) %>% summarise(outgoingRequest = sum(values, na.rm = T))) %>% 
  left_join(data.all[['dublinInR']] %>% group_by(time) %>% summarise(incomingRequest = sum(values, na.rm = T))) %>%
  left_join(data.all[['dublinOutT']] %>% group_by(time) %>% summarise(outgoingTransfer = sum(values, na.rm = T))) %>%
  left_join(data.all[['dublinOutREurodac']] %>% group_by(time) %>% summarise(outgoingEurodacRequest = sum(values, na.rm = T))) %>%
  left_join(data.all[['decisionDublinIn']]) %>%
  left_join(data.all[['decisionDublinOut']]) %>%
  left_join(data.all[['dublinOutEurodacDecision']])

data.all$time <- as.Date(data.all$time)

data.all <- data.all %>% 
  group_by(time) %>% mutate(percOutgoingR = outgoingRequest / values) %>%
  group_by(time) %>% mutate(percOutgoingT = outgoingTransfer / values) %>% 
  group_by(time) %>% mutate(percEurodacOutgoing = outgoingEurodacRequest / outgoingRequest) %>%  
  ungroup()


myAreas <- function(dat = data, x = "time", y = "values", title = "") {
  ggplot(data = dat) + geom_area(aes_string(x, y), fill = "grey") + 
    swi_theme(base_size = 13) + ggtitle(title) +
    scale_x_date(breaks = pretty_breaks(n = 5)) +
    xlab("")  
}

invisible(sapply(1:length(vars), function(i) {
  print(myAreas(dat = data.all, y = vars[i], title = paste0(i, " - ", names(vars)[i])))} 
))


```


### Plot

## Plot general countries' figures 

Interesting charts

* Chart 6: Switzerland seems to have a rather low acceptance rate for Dublin requests
* Chart 7: Switzerland has clearly of the lowest acceptance rate for its Dublin requests (I guess Italy is a special case?)
* Chart 11: The share of eurodac request among all its outgoing Dublin requests is small for Switzerland. 

```{r plot by country, message = F, warning = F, echo = F}
# helper plot
myFacetedAreas <- function(dat = data, x = "time", y = "values", fill = "geo", title = "") {
  ggplot(data = dat) + geom_area(aes_string(x, y, fill = fill)) + 
    facet_wrap( ~ country, ncol = 3) + swi_theme(base_size = 13) + ggtitle(title) +
    scale_x_date(breaks = pretty_breaks(n = 5)) +
    scale_fill_manual(values = swi_rpal) + guides(fill=FALSE) + xlab("")  
}
#myFacetedAreas(title = "Total number of asylum requests")

## Plot


invisible(sapply(1:length(vars), function(i) {
  print(myFacetedAreas(y = vars[i], title = paste0(i, " - ", names(vars)[i])))} 
))
```

## Combine the overall total and subsets in one 

```{r combine}
data.all$country <- data.all$geo <- "TOTAL"
# elegant reordering of columns
data.all %>% select(geo, everything(), country)

data.merged <- rbind(data, data.all)

```